Demo for a Quality of Service networking with Software Defined Network.

### 2017-03-14 report/paper.tex

\setcounter{page}{1}
%\mainmatter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Quality of Service with Software Defined Network.}

@@

\section{Results}\label{sec:results}
\subsection{Controller}
We developed a controller that can handle a mesh network with loops, avoiding
packets flooding. The controller stores the switch and the connection between them and knows which hosts are connected to each switch port starting from the hypothesys that only one
host is connected to each port, unless the port is the one connected to the internet.
With the use of statistics, the controller gets the status of the network and, if necessary,
tries to adjust the traffic and to put in the link with the higher path loss the torrent traffic.
On the other hand it sets up the connection for two gaming hosts going through a link with less delay and tries to redistribute all the 
traffic in a way that all the nodes have an average throughput as lowest as possible.
\\
\newline
We have used some principal functions released with POX controller:
\begin{itemize}
 \item \texttt{pox.openflow.discovery.launch()} this function runs an algorithm in order to discover the network. Every time a new
 topological change occurs this function returns an event on POX with which we can update our topological knowledge.
 It does not provide the host event\cite{pox}.

 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \item \texttt{pox.openflow.spanning_tree.launch()} this function is necessary with a presence of loop on topology
 because with a spanning_tree algorithm it avoid the broadcast storm of message. When we set up \texttt{OFPP_FLOOD} like output port action in a table's
 switch row if some spanning_tree algorithm run the \texttt{OFPP_FLOOD} is only theoretical infact the algorithm close some port for flooding service
 on the switch in order to avoid loop\cite{pox}.
\end{itemize}
\subsection{Real Lab Scenario}
The topological scenario in which we have set up experience is the next:
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/topo.png}
 \caption{\emph{Experience scenario}}
 \label{fig:topo}
\end{figure}

We have use 4 MikroTik Routerboards that were provided us in the Communication's Network Lab at DEI in University of Padua.
The 4 routerboards are connected each other in order to design a loop. Every sigle board connects two host and is connected with a 
legacy switch which provides logically connection with controller.

\subsection{Graph and Network structures}
  In order to achieve the realtime adaptability and statistics storage we developed ad-hoc structure.
  We used a specific Python's library called NetworkX \cite{networkx}. NetworkX allow us to create a graph in witch we can storage proper nodes
  that represent resplectively switchs or hosts elements.
  This topological structure is updated live thanks to connection events and links discover offered by POX's libraries:
  \begin{itemize}
   \item \texttt{_handle_LinkEvent}: from openflow.discovery an event of this type give us two types of information; a new link is added and 
   a still present link is expired;
   \item \texttt{_handle_ConnectionUp}: from core.openflow this event join up when a new switch is added on the network and give us
   some information of the switch, in particular its ports configuration;
   \item \texttt{_handle_ConnectionUp}: from core.openflow like the event explained above comes up when a switch is disconnected from the network;
   \item \texttt{_handle_port_status}: from core.openflow this event occure when on a particular switch there is a change about its ports, for example
   some host is connected or the link is expired;\\
  \end{itemize}

  The link on our network besides has many parameters in witch we can save network's performance. 
  In order to update every links parameters we had to create ad hoc functionality
  because there isn't some function constructed yet.

  This type of structure allow us to run Dijkstra a polynomial algorithm that discover the shorthes path according to various parameters.

  The underlying figure shown the visual output of our laboratory network's topology offered by the controller, we can see 4 different switches and all of the other hosts
  differentiated with their IP address.

@@

  %\#TODO matteo

  \subsection{Traffics analisis}\label{subsec:traf}
  We used laptop connected via
GigabitEthernet interface and Ethernet cable cat.5a to receive the traffic from the destination host.
Our laptop has been connected via Wi-Fi to our access point via protocol 802.11n.
No other traffic sources were active during the testing, so the router was handling only
the traffic from laptop when sniffing traffic. The connection used was an ADSL2+ via copper.
We did 3 type of traffic analysis: Gaming traffic, Skype® and torrent traffic. We analyzed
\#Mb for the gaming traffic, \#Mb for the VoIP traffic and 3GB for the torrent traffic.
The analysis involved the average number of packets per second, the packet size, the protocol (TCP or UDP)

The traffic analysis of the torrent protocol shows that the average packet size is the MTU
(i.e. 1500bytes) and the protocol tryies to exploit all the channel capabilities. The traffic
is made via UDP packets with the port range in the ephimeral ports (2000_- 65535).
Moreover, the node is linked to a lot of IPs with multiple connection: this means that
if someone wants to find a p2p traffic, it can look to the statistics and see if the traffic is loading all the channel,
to how many IPs, an host is connected to and if it uses UDP.
For the VoIP we saw that the protocol uses UDP and the traffic, after a transition time, goes only between two hosts.
The average packet size is the MTU and the protocol tryies to exploit all the channel for a better QoE.
On another test, we tried the VoIP while a download was ongoing, and we saw that the quality
decreased to allow communication. The traffic is between the UDP port 80 \#\#TODO\#\# TOCHECK \#\#

The last traffic analysis we made was with a DOTA[explain what it is] web match.
We saw that for that type of gaming session, the packet size is about \# bytes and involves both TCP and UDP ports.
  \\
  \\
  For every networks performance we have tryed to develop a specific solution. The OpenFlow main goal is to centralize controll and software
  complexity so the routerboard can made only some easy function in order to contain the hardware complexity and mantainly a good speed\cite{qos_paper}.
  Il che ha portato a dover utilizzare stratagemmi.
  \#TODO %riformulare
    \subsubsection{Delay}
Monitoring round-trip time provides important insights for network troubleshooting and traffic engineering.
The  common  monitoring  technique  is  to  actively  send  probe packets  from  selected  vantage  points  (hosts  or  middleboxes).
In traditional  networks, the control over the network routing is
limited,  making  it  impossible  to  monitor  every  selected  path. The  emerging  concept  of  Software  Defined  Networking  sim-plifies
network  control.  However,  OpenFlow,  the  common  SDN protocol,   does   not   support   RTT   monitoring   as   part   of   its
specification. In this experience we have tryed to emply an tecnic to escape this problem\cite{rtt}.
\newline The idea is to measure the time that occure from sending a packet from controller to a switch and the time in which the packet return from
an other switch.
\\
\newline First we have to discovery round trip time from controller to every single switch on our topology.
In order to do this on the controller we start a timer, we send a proper packet to the switch specifing the \texttt{OFPP_CONTROLLER - Send to the controller}
option and when the message return to the controller we stop the timer and calculare RTT for the switch.
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/rtt2.png}
 \caption{\emph{RTT between controller and OF switch}}
 \label{fig:topo}
\end{figure}
\newline
If we need to calculate the delay on a specific link, the rtt has to calculate with the headed switch of this link.
\\
\newline
Now we create a proper message with unique identifier number, we start a timer \texttt{TM1} and then send this message to a first switch specifing the output port.
Then when the packet comes to the second switch it identify the particular message and sends to the controller that stops it's dedicated timer \texttt{TM2}.
\newline This part required a particolar flow entry on the second switch in order to send to the controller only some particulare packet.
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/rtt1.png}
 \caption{\emph{Pck from controller to itself following a particular way.}}
 \label{fig:topo}
\end{figure}
\\
\newline At this point we can calculate the delay of the link.
\\
\\
 delay = (TM2 - TM1) - RTT1/2 - RTT2/2
\\
\\
Teorethically this method must works but in practice there are a lot of delay introduce from computational concestion on the virtualized controller.
What we observe is a estimated delay like some seconds that is impossible, infact if we made some \texttt{ping test} from one host to another host
in this SDN network we realize delay around 3ms.
\\

@@

	\#TODO andrea
	The first data that was collected was the actual link capacity, which depends both on the cable used to connect
	different devices and the effective capacity for each ethernet port linked by the cable itself. Everytime a device supporting
	openflow connects to the controller, it gives to the latter information about its status and its capabilities. One of them is
	the link capacity in three ways: the first is the \textit{supported} capacity, which is the phisical maximum capacity, the second is the
	\textit{advised} capacity, which name describes itself and the latter is the \textit{current} one. We choose to save the last one,
	as it gives the most useful about the link state. This information, moreover, is the one that non-SDN switches use to choose the \textit{metric}
	parameter. The simulation with Mininet showed us that this type of information can be used, but the Routeboards didn't support it.
	As a matter of fact, we weren't able to exploit this capacity in our network realization.
		\\
		\\
  \subsubsection{Throughput}
	\#TODO andrea
	Using the flow statistics, we analyzed the traffic going through a link and, with a simple division, we got the normalized throughput.
	We exploited the information given by the switches with the match structure: this means that we used the throughput for the TCP traffic,
	the one for the UDP and for each of them we could know toward which IP the traffic is directed or is from. This was exploited with mininet
	simulations, but not in the set up in the laboratory. With the latter, we used the statistics as an indicator of the bitrate of a link
	and divided the host to categories based on it.
\\
\subsubsection{Packet error rate}

The port statistics given by the switches contains different informations about the state of each port in a switch. To classify a link for its
reliability, we divided the number of packets with erroneous symbols that couldn't be corrected with a FEC by the total number of packets that were sent,
getting the packet error rate. The choice we made when exploiting this statistic was privileging the gaming and normal traffic, wherease the
heavy traffic was directed to links with highest packet error rate.  As the switches exchanges a packet before recognizing most of the error, both because the
phisical error are corrected using blocks coding and other error are found using a \textit{Frame Check Sequence} like CRC, the byte error were considered worse as
those are indirect measures themself and the error derived from the calculations could grow up.
\#TODO andrea
\\
\subsection{Route changing} %cambio rotte

\#TODO andrea
For each statistic we got and each parameter we obtained, we changed the weight of the graphs. Every graph we used had a specific weight, so the
Dijkstra algorithm for a weighted graph gave different minimum path between two hosts. From the parameter we saved, we choose to change the
route to reach the destination and viceversa. The traffic follow the same path in both the directions and depends only on the type of traffic:
we choose to set the minimum unweighted path if no statistics were available, the minimum throughput path for the standard traffic, the minimum
delay for the gaming one and the maximum packet error rate for the heavy traffic generated by a host.
Two threshold were chosen based on the traffic analysis described on Sec. \ref{subsec:traf} which are 200 bytes/s for the gaming traffic and one megabyte/s
when the node is considered in heavy traffic. Those two threshold were made as they are average values and considering that our network is made of few really fast link,
so on one side the lower limit is sufficient not to consider gaming traffic to be a normal traffic, wherease it's easy to reach the high level of traffic load between two machines.
The changing of the path were made only if the last change happened before one minute because the torrent traffic has, in its protocol, a discovery
algortithm. This means that if a lot of changes were made, most of the traffic generated by the peer to peer, is the control one and the result is that the
load in the network will be heavy and last longer.

\subsection{ARP solution}

During our analysis and in our tests, we prove that in a SDN mesh network, the ARP packets are useless.
This type of packets is useful for a multipoint to multipoint connection where the connection is set between a
HUB and a set of computers or between a bridge and a computer. The first case is no more in use in common networks as
it higher the collision range. The bridges limits their collision range to the port. With openflow each switching unit can
act as a bridge or as a router. In both cases analyses the header of the L2 packet L3 packet and some of the L4 header.
If it has the rule to forward the packet, it will act as declared in the most specific matching rule. Otherwise it will
send to the controller a PacketIn and the controller will act as programmed. If an host want to communicate to another host whithin
the network, it sends the packet to the line and then the switch will check in its tables and execute the actions of the match found.
For an IP network, it will check the path to reach a destination and then send to the defined output port. The mac address in this case
became useless, as the connection is point to point. So, knowing the destination layer2 address is useless as the switch can act
regardless of it. The only thing the switch need to know is the layer2, layer 3 addresses and, at most, the layer 4 address.
This can be seen by a reader as a violation of the layers in the ISO/OSI stack, but to exploit the potentials of the SDN approach,
this is necessary.

\subsection{Default Gateway}
Analyzing the traffic while we were trying the controller on real machines in the lab, we have tried to set the default gateway to
other machines that were connected to the SDN network, but not to the internet. We proved that the controller can redirect correctly the
traffic to the internet. This means that
computers don't need anymore to set a gateway to reach another network as the controller and the switches will redirect the traffic
properly, according to the rules set by the first. The controller must only know which switch ports are connected to the
internet and then add them to the path. I this way, the controller can set a path to reach it and provide connectivity with
a traffic shaping policy and/or differentiate the traffic from and to different networks.

\subsection{Ip switching in pox}

During our tests, both with Mininet and in the lab, we found out that the core of POX, when a PacketIn event occoures, switches the source and the destination addresses,
both MAC and IP, if one of the hosts is already known. We weren't able to understand in which point of the code of POX  and why it happens.
% One of the problem encountered was that, if an host is already found by pox and there is a PacketIn event, the core of the controller
% switches the source and the destination IP and MAC address with the source one.
The problem was partially solved, checking out the
parameters known by the controller, i.e. if the two host are connected to two different switches and the controller switches the source with the
destination address, we can switch it back; the result is that the PacketIn event now has correct informations.
If the two host are in the same switch, the problem is solved using a controlled packet flooding, described in section \ref{sec:connectivity}

\section{Conclusions}\label{sec:conclusion}
Conclusions are a superbrief summary of what has been done and highlighting of the "take home message"
ehi
\#TODO

>>>>

\setcounter{page}{1}
%\mainmatter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Demo for a Quality of Service networking with Software Defined Network.}

@@

\section{Results}\label{sec:results}
\subsection{Controller}
We developed a controller that can handle a mesh network with loops, avoiding
packets flooding. The controller stores the switch and the connection between them and knows which hosts are connected to each switch port starting from the hypothesys that only one
host is connected to each port, unless the port is the one connected to the internet.
With the use of statistics, the controller gets the status of the network and, if necessary,
tries to adjust the traffic and to put in the link with the higher path loss the torrent traffic.
On the other hand it sets up the connection for two gaming hosts going through a link with less delay and tries to redistribute all the 
traffic in a way that all the nodes have an average throughput as lowest as possible.
\\
\newline
We have used some principal functions released with POX controller:
\begin{itemize}
 \item \texttt{pox.openflow.discovery.launch()} this function runs an algorithm in order to discover the network. Every time a new
 topological change occurs this function returns an event on POX with which we can update our topological knowledge.
 It does not provide the host event\cite{pox}.
 \item \texttt{pox.openflow.spanning_tree.launch()} this function is necessary with a presence of loop on topology
 because with a spanning_tree algorithm it avoids the broadcast storm of messages.

When we set up this option \texttt{OFPP_FLOOD} like output port's action in a table's switch row if some spanning_tree algorithm is working
the \texttt{OFPP_FLOOD} is only theoretical in the sense that the algorithm excludes some ports during the flooding option, in order to avoid loop\cite{pox}.
\end{itemize}
\subsection{Real Lab Scenario}
The topological scenario in which we have set up experience is as follows:
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/topo.png}
 \caption{\emph{Experience scenario}}
 \label{fig:topo}
\end{figure}

We have use 4 MikroTik Routerboards that were provided us by the Communication's Network Lab in the department of information  engineering to 
University of Padua.
The 4 routerboards are connected each other in order to design a topological loop. Then every sigle board connects two hosts and is connected with a 
legacy switch which provides logically connection with controller.
The last thing is a host that made gateway's function; we have connected its with both routerboard and legacy network and set up this host like default 
gateway on the controller.

\subsection{Graph and Network's structures}
  In order to achieve the realtime adaptability and statistics storage we developed ad-hoc structures.
  We used a specific Python's library called NetworkX \cite{networkx}. NetworkX allow us to create a graph in witch we can storage suitable nodes
  that represent resplectively switches or hosts.
  This topological structure is updated live thanks to connection events and links discover offered by POX's libraries:
  \begin{itemize}
   \item \texttt{_handle_LinkEvent}: from openflow.discovery an event of this type gives us two types of information; a new link is added and 
   a still present link is expired;
   \item \texttt{_handle_ConnectionUp}: from core.openflow this event joins up when a new switch is added on the network and gives us
   some information of the switch, in particular its ports configuration;
   \item \texttt{_handle_ConnectionUp}: from core.openflow like the event explained above comes up when a switch is disconnected from the network;
   \item \texttt{_handle_port_status}: from core.openflow this event occures when on a particular switch there is a change about its ports, for example
   a new host has just been connected or the link is expired;\\
  \end{itemize}

  The links on our network logically have many parameters in witch we can storage network's performance; in practice a NetworkX's graph support only 
  one or two link's attributes so we have create a lots of graphs with the same topological information but different link's attributes.

  In order to update every link's parameters we had to create ad-hoc functionality
  since there are not proper functions constructed before.

  This type of structure allow us to run Dijkstra's polynomial algorithm that discover the shortest path according to various parameters.

  The following picture shows the visual output of our laboratory network's topology offered by the controller, we can see 4 different switches and all of the other hosts
  differentiated with their IP addresses.

@@

\subsection{Traffic's Analysis}\label{subsec:traf}
We used laptop connected via
GigabitEthernet interface and Ethernet cable cat.5a to receive the traffic from the destination host.
Our laptop has been connected via Wi-Fi to our access point via protocol 802.11n.
No other traffic sources were active during the testing, so the router was handling only
the traffic from laptop when sniffing traffic. The connection used was an ADSL2+ via copper.
We did 3 type of traffic analysis: Gaming traffic, Skype® and torrent traffic. We analyzed
\#Mb for the gaming traffic, \#Mb for the VoIP traffic and 3GB for the torrent traffic.
The analysis involved the average number of packets per second, the packet size, the protocol (TCP or UDP).

The traffic analysis of the torrent protocol shows that the average packet size is the MTU
(i.e. 1500bytes) and the protocol tryies to exploit all the channel capabilities. The traffic
is made via UDP packets with the port range in the ephimeral ports (2000_- 65535).
Moreover, the node is linked to a lot of IPs with multiple connection: this means that
if someone wants to find a p2p traffic, it can look to the statistics and see if the traffic is loading all the channel,
to how many IPs, an host is connected to and if it uses UDP.
For the VoIP we saw that the protocol uses UDP and the traffic, after a transition time, goes only between two hosts.
The average packet size is the MTU and the protocol tryies to exploit all the channel for a better QoE.
On another test, we tried the VoIP while a download was ongoing, and we saw that the quality
decreased to allow communication. The traffic is between the UDP port 80.

  For every networks performance we have tryed to develop a specific solution. The OpenFlow main goal is to centralize controll and software
  complexity so the routerboards can made only some easy function in order to contain the hardware complexity and mantainly a good speed\cite{qos_paper}.
  This led to the necessity of stratagems utilization.

\subsubsection{Delay}
Monitoring round-trip time provides important insights for network troubleshooting and traffic engineering.
The  common  monitoring  technique  is  to  actively  send  probe packets  from  selected  vantage  points  hosts  or  middleboxes.
In traditional  networks, the control over the network routing is
limited,  making  it  impossible  to  monitor  every  selected  path. The  emerging  concept  of  Software  Defined  Networking  simplifies
network  control.  However,  OpenFlow,  the  common  SDN protocol,   does   not   support   RTT   monitoring   as   part   of   its
specification. In this experience we have tryed to emply an tecnic to escape this problem\cite{rtt}.
\newline The idea is to measure the time that occure from sending a packet from controller to a switch and the time in which the packet return from
an other switch.
\\
\newline First we have to discovery round trip time from controller to every single switch on our topology.
In order to do this on the controller we start a timer, we send a proper packet to the switch specifing the \texttt{OFPP_CONTROLLER - Send to the controller}
option and when the message return to the controller we stop the timer and calculare RTT for the switch.
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/rtt2.png}
 \caption{\emph{RTT between controller and OF switch}}
 \label{fig:topo}
\end{figure}
\newline
If we need to calculate the delay on a specific link, the rtt has to calculate with the headed switch of this link.
\\
\newline
Now we create a proper message with unique id. number, let's start a timer \texttt{TM1} and then send this message to a first switch specifing the output port.
Then when the packet comes to the second switch it identify the particular message and sends to the controller that stops it's dedicated timer \texttt{TM2}.
\newline This part required a particolar flow entry on the second switch in order to send to the controller only some particulare packet.
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/rtt1.png}
 \caption{\emph{Pck from controller to itself following a particular way.}}
 \label{fig:topo}
\end{figure}
\\
\newline At this point we can calculate the delay of the link.
\\
\\
 delay = (TM2 - TM1) - RTT1/2 - RTT2/2
\\
\\
Teorethically this method must works but in practice there are a lot of delay introduced from computational congestion on the virtualized controller.
What we observe is a estimated delay like a few seconds which is not the case, infact if we made some \texttt{ping test} from one host to another host
in this SDN network we realize delay around 3ms.
\\

@@

	The first data that was collected was the actual link capacity, which depends both on the cable used to connect
	different devices and the effective capacity for each ethernet port linked by the cable itself. Everytime a device supporting
	openflow connects to the controller, it gives to the latter information about its status and its capabilities. One of them is
	the link capacity in three ways: the first is the \textit{supported} capacity, which is the phisical maximum capacity, the second is the
	\textit{advised} capacity, which name describes itself and the latter is the \textit{current} one. We choose to save the last one,
	as it gives the most useful about the link state. This information, moreover, is the one that non-SDN switches use to choose the \textit{metric}
	parameter. The simulation with Mininet showed us that this type of information can be used, but the Routeboards didn't support it.
	As a matter of fact, we weren't able to exploit this capacity in our network realization.
		\\
		\\
  \subsubsection{Throughput}

	Using the flow statistics, we analyzed the traffic going through a link and, with a simple division, we got the normalized throughput.
	We exploited the information given by the switches with the match structure: this means that we used the throughput for the TCP traffic,
	the one for the UDP and for each of them we could know toward which IP the traffic is directed or is from. This was exploited with mininet
	simulations, but not in the set up in the laboratory. With the latter, we used the statistics as an indicator of the bitrate of a link
	and divided the host to categories based on it.
\\
\subsubsection{Packet error rate}
The port statistics given by the switches contains different informations about the state of each port in a switch. To classify a link for its
reliability, we divided the number of packets with erroneous symbols that couldn't be corrected with a FEC by the total number of packets that were sent,
getting the packet error rate. The choice we made when exploiting this statistic was privileging the gaming and normal traffic, wherease the
heavy traffic was directed to links with highest packet error rate.  As the switches exchanges a packet before recognizing most of the error, both because the
phisical error are corrected using blocks coding and other error are found using a \textit{Frame Check Sequence} like CRC, the byte error were considered worse as
those are indirect measures themself and the error derived from the calculations could grow up.

\subsection{Route changing}

For each statistic we got and each parameter we obtained, we changed the weight of the graphs. Every graph we used had a specific weight, so the
Dijkstra algorithm for a weighted graph gave different minimum path between two hosts. From the parameter we saved, we choose to change the
route to reach the destination and viceversa. The traffic follow the same path in both the directions and depends only on the type of traffic:
we choose to set the minimum unweighted path if no statistics were available, the minimum throughput path for the standard traffic, the minimum
delay for the gaming one and the maximum packet error rate for the heavy traffic generated by a host.
Two threshold were chosen based on the traffic analysis described on Sec. \ref{subsec:traf} which are 200 bytes/s for the gaming traffic and one megabyte/s
when the node is considered in heavy traffic. Those two threshold were made as they are average values and considering that our network is made of few really fast link,
so on one side the lower limit is sufficient not to consider gaming traffic to be a normal traffic, wherease it's easy to reach the high level of traffic load between two machines.
The changing of the path were made only if the last change happened before one minute because the torrent traffic has, in its protocol, a discovery
algortithm. This means that if a lot of changes were made, most of the traffic generated by the peer to peer, is the control one and the result is that the
load in the network will be heavy and last longer.

\subsection{ARP solution}
During our analysis and in our tests, we prove that in a SDN mesh network, the ARP packets are useless.
This type of packets is useful for a multipoint to multipoint connection where the connection is set between a
HUB and a set of computers or between a bridge and a computer. The first case is no more in use in common networks as
it higher the collision range. The bridges limits their collision range to the port. With openflow each switching unit can
act as a bridge or as a router. In both cases analyses the header of the L2 packet L3 packet and some of the L4 header.
If it has the rule to forward the packet, it will act as declared in the most specific matching rule. Otherwise it will
send to the controller a PacketIn and the controller will act as programmed. If an host want to communicate to another host whithin
the network, it sends the packet to the line and then the switch will check in its tables and execute the actions of the match found.
For an IP network, it will check the path to reach a destination and then send to the defined output port. The mac address in this case
became useless, as the connection is point to point. So, knowing the destination layer2 address is useless as the switch can act
regardless of it. The only thing the switch need to know is the layer2, layer 3 addresses and, at most, the layer 4 address.
This can be seen by a reader as a violation of the layers in the ISO/OSI stack, but to exploit the potentials of the SDN approach,
this is necessary.

\subsection{Default Gateway}

Analyzing the traffic while we were trying the controller on real machines in the lab, we have tried to set the default gateway to
other machines that were connected to the SDN network, but not to the internet. We proved that the controller can redirect correctly the
traffic to the internet. This means that
computers don't need anymore to set a gateway to reach another network as the controller and the switches will redirect the traffic
properly, according to the rules set by the first. The controller must only know which switch ports are connected to the
internet and then add them to the path. I this way, the controller can set a path to reach it and provide connectivity with
a traffic shaping policy and/or differentiate the traffic from and to different networks.

\subsection{Ip switching in pox}
During our tests, both with Mininet and in the lab, we found out that the core of POX, when a PacketIn event occoures, switches the source and the destination addresses,
both MAC and IP, if one of the hosts is already known. We weren't able to understand in which point of the code of POX  and why it happens.
% One of the problem encountered was that, if an host is already found by pox and there is a PacketIn event, the core of the controller
% switches the source and the destination IP and MAC address with the source one.
The problem was partially solved, checking out the
parameters known by the controller, i.e. if the two host are connected to two different switches and the controller switches the source with the
destination address, we can switch it back; the result is that the PacketIn event now has correct informations.
If the two host are in the same switch, the problem is solved using a controlled packet flooding, described in section \ref{sec:connectivity}.

\section{Conclusions}\label{sec:conclusion}
% Conclusions are a superbrief summary of what has been done and highlighting of the "take home message"
% ehi
For this course project, the report shows how the demo was made, in particular what problem we encountered
and what we exploited, showing what protocols could be unused in the future. We showed a possible simple data structure
that can allow the controller to be able to know the network, its topology and which route to set, based on the traffic.
Morover, we showed how to exploit the potentialities of the SDN network, giving particular importance on the
statistics and the path switching capabilities.
The multi-graph solution allowed us to connect the network and, based on its condition, choose which weight to apply
in order to give a Quality of Service based network. With some boards that supports the newer version of Openflow
(i.e. 1.5, the latest), the code can be ported, so it can be used with the Ryu core controller and more statistic can
be exploited. Moreover, the controller can set more strictive rules for the flow control and can get more informations from the
switches.

% mi pare vada bene, l ultima versione di pox dovrebbe essere almeno la 1.5.1
